---
title: "Exam 2 Submission"
author: "Echo Nattinger"
date: "6/28/2021"
output: word_document
---

## College Scorecard Questions

```{r setup}
# Clearing the environment and loading the dataset
rm(list=ls(all=TRUE))
library(rio)
college_scorecard = import("2021_exam2_data.xlsx", which=4)
# Providing summary statistics
summary(college_scorecard)
```

```{r}
# Creating smaller dataset
library(tidyverse)
small_scorecard = college_scorecard%>%
  filter(year == 2014 | year == 2015)%>%
  filter(state_abbr == "TX" | state_abbr == "LA")
  
```

```{r}
# Collapsing into smaller dataframe
# First have to get rid of NAs in count_working
small_scorecard_no_NA = small_scorecard %>% na.omit(small_scorecard$count_working)
even_smaller_scorecard = small_scorecard_no_NA %>%
  group_by(state_abbr)%>%
  summarize(avwork = mean(count_working), 
            totalworking = sum(count_working))

# Graphing
library(ggplot2)

ggplot(even_smaller_scorecard, aes(state_abbr, totalworking/sum(totalworking)))+
  geom_col(stat="identify")+
  labs(title="Share of Employment by Employee's College State", 
       x="State", 
       y="Percentage of Total Employment")

summary(even_smaller_scorecard)
table(small_scorecard_no_NA$state_abbr)

```
On the basis of this graph, people who graduated from four-year colleges/universities in Texas had a better change of being employed. I do think going to college in certain states could give a student a better chance of employment, but I don't think that's what this graph proves. The average number of people working from each state is relatively close (2479 and 2207 for LA and TX respectively). Further, TX has way more schools in this dataset than LA.

## Avocado Sales
```{r}
# Loading dataset
avocados = import("2021_exam2_data.xlsx", which=2)
```

```{r}
# Creating new variable
library(lubridate)
avocados = avocados%>%
  mutate(year = lubridate::year(avocados$date))
```

```{r}
# Deflating
library(WDI)
# loading deflator data and renaming it for ease
deflatordata = WDI(country="US", indicator=c("NY.GDP.DEFL.ZS"), 
                   start=1960, end=2018, extra=FALSE, cache=NULL)
library(data.table)
setnames(deflatordata, "NY.GDP.DEFL.ZS", "deflator")
deflatordata$country = NULL
deflatordata$iso2c = NULL
# Merging
deflatedavocados = left_join(avocados, deflatordata, by=c("year"))

# Creating deflated amounts
deflatedavocados$deflated_price_2015 = deflatedavocados$average_price/(deflatedavocados$deflator/100)

```

```{r}
# Collapsing deflatedavocados
collapsed_avocados = deflatedavocados%>%
  group_by(year)%>%
  summarize(avdeflatedprice = mean(deflated_price_2015))

head(collapsed_avocados)
```

```{r}
# Reshaping the data
wide_avocados = collapsed_avocados%>%
  pivot_wider(id_cols = "year", 
              names_from = "year", 
              values_from = "avdeflatedprice" )
head(wide_avocados)
```

```{r}
# Labeling variables in wide_avocados
library(labelled)
var_label(wide_avocados) = list(`2015` = "Year 2015", 
                                `2016` = "Year 2016", 
                                `2017` = "Year 2017", 
                                `2018` = "Year 2018")
```

## Training Dataset
```{r}
# Loading dataset
training = import("2021_exam2_data.xlsx", which=3)
# Creating ID variable
training$id = 1:nrow(training)
# Reshaping
traininglong = training%>%
  pivot_longer(cols = starts_with("re"), 
              names_to = NULL, 
             values_to= "earnings",
             values_drop_na = FALSE)
# Summary
summary(traininglong)
```

## Titanic Questions
```{r}
# Loading dataset and providing summary statistics
titanic = import("2021_exam2_data.xlsx", which=1)
summary(titanic)
```

```{r}
# Crosstabs
library(doBy)
summaryBy(survived ~ female, data = titanic, FUN=mean)
```

The results of this cross-tab show that those who are female (female=1) were less likely than their male counterparts to survive (with a mean of 0.21 compared to 0.73 for men).

```{r}
# If else and frequency table
titanic$first_class = ifelse(titanic$class == 1, 1, 0)
table(titanic$first_class)
```

Bonus: "My Heart Will Go On" by Celine Dion

Github link: <https://github.com/enatt/DataSciExam2>